Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-17 chi2012.tex
minor edits

The two prestudies provided valuable insight towards this goal that allowed us to guide our design process. The first prestudy confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.  The second prestudy revealed that the most common inconisistencies that occurred in the APIs and data feeds available through from such data sources consisted predominantly of the simpler kinds of heterogeneity - terminological and simpler forms of structural heterogeneity, rather than the delicate semantic/modeling variety, which were comparatively rare. This gave us optimism that an interactional approach optimised around the these simpler kinds of heterogeneity would be effective.
[0
In the following sections we describe the design process we used to create prototype DataPalette, followed by a detailed description of the design of the prototype and justification.

@@

\item Supporting dereferencing along different paths \emph{by structure} allows instances with different property names, and even sub-structures to be quickly consolidated by separate dereferencing, thereby effectively coping with terminological and structural hetereogeneity.
\end{enumerate}

@@

\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=18cm]{img/sameas}
\caption{\emph{Coreference consolidation with drag-and-drop Same-As} - When two instances represent the same entity, users can drag one on top of the other, and they are combined - effectively displaying the union of properties of both source entities. This can be un-done by deleting the Same-As relaitonship in the view in the lower-left corner of the workspace.}
\label{fig:sameas}
\end{center}
\end{figure*}

@@

Related to the concept of co-reference, PS.2 revealed that enumerated values of properties were often inconsistent, e.g. ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property. To reconcile and consolidate corresponding values from diffeent  datasources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.

@@

\subsubsection{Study and task design}
To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two predefined tasks in separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'' which we describe below).  We used upon the within-subjects A-B format in order to make both quantiative measurements about an individual's performance with DP (relative to baseline), but also to make qualitative observations of the processes used by each participant as they solved each task.  Each task was paired with one of two datasets, depending on condition, so that we could evaluate whether the participant's performance of completing a task differed with different datasets.  Conditions and interface order were fully counterbalanced among participants to eliminate ordering and task bias.

@@

\subsection{Condition B (control}
In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.  Participants who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues. In particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of university teaching by the Times Higher Education. There were numerous usability issues in using websites, particularly in searching for university courses and course requirements. Similarly, participants found it difficult to understand the locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto the same map using the websites. Overall, the users that spent the most time using websites exhibited significant frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations), and all ended up falling back to using the supplied spreadsheets.

>>>>

The two pre-studies provided valuable insight towards this goal that allowed us to guide our design process. The first pre-study confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.  The second pre-study revealed that the most common inconsistencies that occurred in the APIs and data feeds available through from such data sources consisted predominantly of the simpler kinds of heterogeneity - terminological and simpler forms of structural heterogeneity, rather than the delicate semantic/modelling variety, which were comparatively rare. This gave us optimism that an interactional approach optimised around the these simpler kinds of heterogeneity would be effective.
[0
In the following sections we describe the design process we used to create prototype DataPalette, followed by a detailed description of the design of the prototype and justification.

@@

\item Supporting dereferencing along different paths \emph{by structure} allows instances with different property names, and even sub-structures to be quickly consolidated by separate dereferencing, thereby effectively coping with terminological and structural heterogeneity.
\end{enumerate}

@@

\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=18cm]{img/sameas}
\caption{\emph{Coreference consolidation with drag-and-drop Same-As} - When two instances represent the same entity, users can drag one on top of the other, and they are combined - effectively displaying the union of properties of both source entities. This can be un-done by deleting the Same-As relationship in the view in the lower-left corner of the workspace.}
\label{fig:sameas}
\end{center}
\end{figure*}

@@

Related to the concept of co-reference, PS.2 revealed that enumerated values of properties were often inconsistent, e.g. ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property. To reconcile and consolidate corresponding values from different  datasources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.

@@

\subsubsection{Study and task design}
To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two predefined tasks in separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'' which we describe below).  We used upon the within-subjects A-B format in order to make both quantitative measurements about an individual's performance with DP (relative to baseline), but also to make qualitative observations of the processes used by each participant as they solved each task.  Each task was paired with one of two datasets, depending on condition, so that we could evaluate whether the participant's performance of completing a task differed with different datasets.  Conditions and interface order were fully counterbalanced among participants to eliminate ordering and task bias.

@@

\subsection{Condition B (control)}
In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.  Participants who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues. In particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of university teaching by the Times Higher Education. There were numerous usability issues in using websites, particularly in searching for university courses and course requirements. Similarly, participants found it difficult to understand the locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto the same map using the websites. Overall, the users that spent the most time using websites exhibited significant frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations), and all ended up falling back to using the supplied spreadsheets.
